A Plasma Torus Around a Young Low-Mass Star

### 2025-03-24 paper/ms.tex
update

{\it Age: Isochrones}---The color and absolute magnitude of
TIC~141146667 suggest that it is a pre-main sequence M dwarf, similar to
all other known CPVs \cite{Stauffer2017,Stauffer2021,Bouma2024}.  The
star's proximity ($d$=58\,pc) and high galactic latitude
($b$=$+$53$^\circ$) yield negligible interstellar reddening along the
line of sight \cite{Green2019}.  Figure~\ref{fig:camd} shows the
location of the star in the color--absolute magnitude diagram (CAMD)
relative to young stellar populations including Upper Scorpius (USco),
IC~2602, and the Pleiades.  To make this diagram, we adopted the USco
members in the $\delta$~Sco and $\sigma$~Sco sub-associations from
\cite{Ratzenbock2023}, and the IC~2602 and Pleiades members from
\cite{Hunt2024}.  We assumed an average $V$-band extinction $A_{\rm
V}$=$\{0.12, 0.11, 0.10\}\,{\rm mag}$ for USco \cite{Pecaut2016},
IC~2602 \cite{Hunt2024}, and the Pleiades \cite{Hunt2024} respectively.
We dereddened the photometry using the extinction coefficients
$k_X\equiv A_X/A_0$ tabulated in \cite{GaiaCollaboration2018}, assuming
that $A_0 = 3.1 E(B-V)$.

Figure~\ref{fig:camd} shows that TIC~141146667 falls between the USco
and Pleiades sequences, while nearly overlapping with IC~2602.  However,
the precision of the implied age is set by the intrinsic scatter of
these calibration sequences; the most luminous stars in the Pleiades of
the same color have a similar absolute magnitude as TIC~141146667.
Previous work \cite{Stauffer2021} has also noted that in the Gaia
passbands, CPVs tend to be photometrically redder and more luminous than
single stars in any given cluster, similar to other rapid rotators.
While this effect complicates any attempt at age inference based on the
Gaia photometry, it suggests that the Pleiades may be a better
comparison population than IC~2602.  We take the star's
location in the color--absolute magnitude diagram to suggest age bounds
$t_{\rm CAMD}$$\sim$30-150\,Myr.

@@

Finally, a constraint on the magnetic field strength at the site of
the clump follows from the requirement that the magnetic pressure
exceed the thermal pressure, $B^2 / 8\pi > n_{\rm H} k T$.  Although
we do not know the plasma temperature, if it were significantly beyond
1,000-10,000\,K, we would either fully ionize the hydrogen, or not
ionize enough of it.  We can therefore estimate
\begin{equation}
  B \gtrsim 3\,{\rm G}
  \left(
  \frac{n_{\rm H}}{5\times10^{11}\,{\rm cm}^{-3}}
  \frac{T}{5000\,{\rm K}}
  \right)^{1/2}.
\end{equation}
Given that the average surface magnetic field strengths of low-mass
stars have been measured to span hundreds to thousands of Gauss
\cite{Donati2009,Kochukhov2021,Reiners2022}, this bound is easily met.

>>>>

{\it Age: Isochrones}---The color and absolute magnitude of
TIC~141146667 suggest that it is a pre-main sequence M dwarf, similar to
all other known CPVs \cite{Stauffer2017,Stauffer2021,Bouma2024}.  The
star's proximity ($d$=58\,pc) and high galactic latitude
($b$=$+$53$^\circ$) yield negligible interstellar reddening along the
line of sight \cite{Green2019}.  Figure~\ref{fig:camd} shows the
location of TIC~141146667 in the color--absolute magnitude diagram
(CAMD) relative to young stellar populations including Upper Scorpius
(USco), IC~2602, and the Pleiades.  To make this diagram, we adopted
the USco members in the $\delta$~Sco and $\sigma$~Sco sub-associations
from \cite{Ratzenbock2023}, and the IC~2602 and Pleiades members from
\cite{Hunt2024}.  We assumed an average $V$-band extinction $A_{\rm
V}$=$\{0.12, 0.11, 0.10\}\,{\rm mag}$ for USco \cite{Pecaut2016},
IC~2602 \cite{Hunt2024}, and the Pleiades \cite{Hunt2024}
respectively, and ages of 8\,Myr \cite{Ratzenbock2023}, 40\,Myr
\cite{Randich2018}, and 112\,Myr \cite{Dahm2015} for each respective
cluster.  We dereddened the photometry using the extinction
coefficients $k_X\equiv A_X/A_0$ tabulated in
\cite{GaiaCollaboration2018}, assuming that $A_0 = 3.1 E(B-V)$.

Figure~\ref{fig:camd} shows that TIC~141146667 falls between the USco
and Pleiades sequences, and approximately overlaps IC~2602.  However,
the precision of the implied age is set by the intrinsic scatter of
these calibration sequences; the most luminous stars in the Pleiades of
the same color have a similar absolute magnitude as TIC~141146667.
Previous work \cite{Stauffer2021} has also noted that in the Gaia
passbands, CPVs tend to be photometrically redder and more luminous than
single stars in any given cluster, similar to other rapid rotators.
While this effect complicates any attempt at age inference based on the
Gaia photometry, it suggests that the Pleiades may be a better
comparison population than IC~2602.  We take the star's
location in the color--absolute magnitude diagram to suggest age bounds
$t_{\rm CAMD}$$\sim$30-150\,Myr.

@@

Finally, a constraint on the magnetic field strength at the site of
the clump follows from the requirement that the magnetic pressure
exceed the thermal pressure, $B_{\rm c}^2 / 8\pi > n_{\rm H} k T$.  Although
we do not know the plasma temperature, if it were significantly beyond
1,000-10,000\,K, we would either fully ionize the hydrogen, or not
ionize enough of it.  The field strength at the clump must therefore
exceed
\begin{equation}
  B_{\rm c} \gtrsim 3\,{\rm G}
  \left(
  \frac{n_{\rm H}}{5\times10^{11}\,{\rm cm}^{-3}}
  \frac{T}{5000\,{\rm K}}
  \right)^{1/2}.
\end{equation}
Given that the average surface magnetic field strengths of low-mass
stars have been measured to span hundreds to thousands of Gauss
\cite{Donati2009,Kochukhov2021,Reiners2022}, this bound is easily met.
